### 2014-03-01 paper.tex
changed abstract

%% bare_jrnl.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
   \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
\usepackage{amsfonts}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize,caption=false]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Community-Centric Innovation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{
    Jaume~Barcelo~%\IEEEmembership{Member,~IEEE,}
    and~Boris~Bellalta%~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\thanks{The authors are with Universitat Pompeu Fabra.\newline
Address: Roc Boronat 138, 08018, Barcelona. \newline
E-mail: {name.surname}@upf.edu }%
}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
The Internet is profoundly transforming the economic and social structures of the 21st century.
The ability to instantaneously share data across the globe enables new ways of collaboration and new production models.
The individuals can join loosely organized working groups and communities and contribute to the evolution and progress of the humankind.
By choosing which communities they join and which communities they contribute to, the users can also be participants and producers.
%Furthermore, this model enables a new way of democracy in which the people vote not only with their ballots, but also with their daily jobs and contributions.
In this work we describe examples of community-centric innovation in which the users play a central role in the development of the project.
The emphasis is placed in community networks and networked cryptocurrencies.
These technologies are good examples in the sense that they empower the communities that use and develop them.
These are initiatives by and for the people.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
 bottom-up broadband, community networks
%Slotted Aloha, game theory, contention control, media access control.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{he} growth of the Internet has dramatically accelerated the path of technological evolution.
The amount of information and knowledge available on the Net is increasing.
Importantly,  the number of people that can access that knowledge is increasing as well.
The Internet is more than an incremental evolution that brings faster progress.
It is a disruptive innovation that is completely transforming the way we organize ourselves and relate to each other.

The fact that the information can be shared with little restrictions in terms of space, time and cost, makes it possible for people to organize in horizontal structures.
In this horizontal models, also known as peer-to-peer models, all the participants can access to the information required to contribute to the decision-making process.
By removing the barriers for participation, the working force of peer-to-peer projects is reinforced.
The benefit is not only the number of participants, but also the variety.
The heterogeneous background of the participants makes it possible to consider a wider space of solutions and the extensive scrutiny throughout the discussion and decision-making stage reduces the chances of mistakes.

In this work we consider several examples in which the users organize in communities to construct user-centric networks and networked services.
We start with community networks in Sec.~\ref{sec:community-networks} and then continue with networked cryptocurrencies in Sec.~\ref{sec:networked-cryptocurrencies}.
Other technologies fruit of community-centric innovation are outlined in Sec.~\ref{sec:other} and an overview of research and innovation projects related to community-centric innovation is presented in Sec.~\ref{sec:projects}.
Finally, we conclude in Sec.~\ref{sec:conclusion}.

\section{Community Networks}
\label{sec:community-networks}

Probably the best example of user-centric networking are community networks.
Community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the users' needs.
This needs are not necessarily only about connectivity.
The users might be interested in learning about the network, and in researching  new services, applications and protocols.

There are cases in which the participants of community networks include not only individuals, but also local organizations such as municipalities or universities \cite{bogliolo2007uwc}.
This involvement stems from the fact that community network have been effective in overcoming the digital divide \cite{oliver2010wca}.

Community networks are particularly useful in underserved areas with little presence of commercial operators.
In rural, distant or poor areas community networks can be the only viable alternative to access the Internet.
In these places, the inhabitants have a strong incentive to spend time and energies in acquiring the necessary skills required to deploy and operate a network.

Another aspect that can draw the attention to community networks is privacy.
As commercial networks are involved in mass surveillance programs, the citizens see in community networks a way to opt out from privacy-invading campaigns.
Finally, community networks are also a means of getting around censorship.
Totalitarian regimes build censorship into the networks to control which information is distributed among the population.
Filtering the information and Internet blackouts have been used as ways for hindering social movements and placate population unrest.

Compared to commercial Internet, Community Networks are more heterogeneous and decentralized.
As a consequence they have properties that complement those of commercial networks.
In case of natural disaster, having alternative resilient networks to coordinate rescue and humanitarian help can save lives.

Mesh technologies such as the \emph{quick mesh project} (qmp) are evolving to make it easier for regular citizens to set up a network in a matter of minutes.
Friendly configuration web interfaces can be used to abstract the underlying technological complexity.
In community networks, local people become familiar with the procedures required to set up, maintain and troubleshoot the network.
In case of emergency, the presence of local experts facilitates quick deployments and reparations.

The decentralized nature of community networks greatly helps in fostering organic growth and also in  resisting node failures.
The absence of central points of failure guarantees that two neighbours will be able to connect to each other as long as they keep their nodes running, regardless of what happens to the rest of the network.
Mesh routing protocols, such as BMX6 \cite{neumann2012ebc}, allow the network to quickly re-configure itself and to route packets around failures to preserve connectivity.
These protocols are particularly tailored to work in challenging multi-hop wireless conditions.
To take down a community network it is necessary to shut down all the nodes.

Community networks are often less stable and more dynamic than their commercial counterparts.
The participants of community networks are eager to experiment and evolve all aspects of the network even if this results in an increased risk of occasional connectivity cuts.
As an example, while commercial networks run mostly on IPv4, qmp networks are native IPv6 networks.
Protocols, firmware and topology change on a daily basis to fix bugs, improve the performance or adapt to new situations.

This continuous change that forces to keep the equipment's firmware updated represents  an additional layer of protection against spying agencies interested in compromising the network.
It is much more difficult to attack up-to-date systems than systems running old unpatched firmwares.

Beyond the purely technological aspects, the importance of community networks is precisely the fact that they are governed by a community.
Community governance is drastically different from hierarchical governance.
For the network to hold together, it is necessary that the participants reach some kind of consensus at the time of making decisions.
It is not uncommon to see different participants or groups of participants having completely different interests, ethical values or methodologies.
Still, it is in the benefit of all the participants to reach a consensus and agree in some basic principles.
The continuous negotiation, discussion and clashes of completely different positions can reach a level that threatens the integrity of the network.

The participation in a community network trains the citizens in self-management techniques that inherit many of the properties of decentralized networks.
The lack of a central authority obliges the participants to take responsibility, exercise their freedom and respect the others.
Making concessions is necessary in order to achieve common goals.
%The organizational structures that we find in community networks mimic those of free (libre) software communities and other Internet-based coordinated projects such as Wikipedia.

All participants are invited to contribute, and contributions are evaluated in an informal process for merit or weaknesses.
The participants can progressively build up a reputation by making high-quality contributions and helping others to accomplish their goals.
The organization progressively evolves towards a meritocratic systems in which those participants that consistently contribute to the value of the community earn the respect of their peers and emerge as natural leaders.
In any case, this is not a leadership which is attached to the person. 
The leadership is attached to the contributions.

A core principle of community networks is precisely the sharing of the network as a commons good.
This is particularly relevant in the case of wireless networks that use the ISM (industrial, scientific, medical) bands.
This bands contain a small fraction of the spectrum that is available without requiring a licence. 
In practice, these bands are a commons good that should be carefully managed by the community to use it effectively without creating unnecessary interference.
Good practices include the use of directive antennas and the limitation of the radiated power.

As the community network is a commons resource, there is no single owner of the network.
%In fact there is no owner of the network.
Even though the participants may own the pieces of hardware that they contributed to the network, the network as a whole belongs the community.
As the value of the network increases with the number of participants, the community has an incentive to share existing network resources with newcomers.

By being part of the community network, the participants agree in altruistically routing traffic for other participants.
Besides connectivity services, the community networks can also offer networked services such as email, mailing lists, HTTP proxy services, FTP, collaborative documents and social platforms.

There is a large number of community networks and they are very heterogeneous \cite{avonts2013qbe}.
They use different technologies such as wireless point-to-point links, wireless multi-point to multi-point links, ADSL or fiber.
Even in the case in which two community networks use the same layer-2 technology, it is still likely that they use different hardware, firmware or protocols.
The sizes of community networks also vary greatly and can range from few tenths of nodes to tenths of thousands.
The amount of documentation, supporting applications and organizational structure is also very different in different community networks.

Despite the disparity of models for community networks, there is an aspect that is common to all of them.
All these network are user-centric.
They are planned, financed, deployed and maintained by the same users that benefit from them.

\section{Networked Cryptocurrencies}
\label{sec:networked-cryptocurrencies}

Another area that has witnessed the potential for community-centric innovation is that of networked cryptocurrency \cite{nakamoto2008bpe}.
The combination of cryptography, free libre software and a data network makes it possible the use of networked cryptocurrencies.
In these technologies, a cryptographically secure distributed ledger is maintained by the network.
The network is composed of nodes that interact in a peer-to-peer fashion following a common set of rules (i.e., a protocol) to guarantee the integrity of the cryptocurrency.
The network not only distributes information about the ongoing transactions, it also provides hashing power that is used as a proof of work.
The distributed ledger is protected by the proof of work.
In a sense, the nodes vote with their computational work and the majority decides which transactions are included in the public distributed ledger.

Cryptocurrency has some advantages when compared to traditional fiat currencies.
Some of the advantages of cryptocurrencies are related to their peer-to-peer nature.
One of them is the lack of a central authority.
A central authority can make arbitrary decisions and create money at no cost.
In cryptocurrencies, the emission of money is controlled by the protocol and governed by publicly known rules.
This means that the availability of money at any point in the future is known by all the participants.
The emitted money is used to reward the nodes that contribute with hashing power to the security of the network.
Hashing power requires hardware and electricity, which means that it is no longer possible to create money at zero cost.
In fact, the money generation business (popularly known as \emph{mining}) is, in a financial sense, a risky endeavour.
The cost of hashing power required to \emph{mine} can exceed the value of the generated money.

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
The risks of blockades or political interference are greatly mitigated.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency.
In some cases the reason to bypass government or paying networks blockades and in others the motivation is to operate in countries with lacking payment networks.

It is in areas lacking effective financial and banking systems where cryptocurrencies have a potential for greater impact.
Just like some areas leapfrogged from no telecommunication services to wireless handheld Internet enabled devices, it is possible that these same areas that currently use mobile calls minutes as a currency transition directly to the latest currency technology which is cryptocurrency.
Many countries suffer from chronic hyperinflation problems that could be solved by moving to a cryptocurrency in which the inflation rate (if any) can be chosen by the community and embedded in the protocol.

Some of the advantages of networked cryptocurrencies are derived from their networked nature.
The reach of the cryptocurrency is the same as that of the underlying network.
If a cryptocurrency runs on the Internet, the scope is global.
As a result, economic transactions benefit from all the advantages of Internet protocols.
Sending money is as simple as sending an electronic mail.
A cryptocurrency transaction propagates almost instantaneously across all the Internet and is confirmed in a matter of minutes.
Payments can be easily automatized using scripts or other software.
Any networked device, such as servers, laptops or mobile phones, can be used to send a payment.
Furthermore, the payment network is available 24 hours a day and 365 days a year.

There are also properties of networked cryptocurrencies related to their cryptographical nature.
These currencies do not require the users to trust a third party.
The trust relies on cryptographic primitives that have proven to be robust over the years.
The reliance on secret keys makes it possible to save the same money in more than one place at the same time.
Even if one of the copies of the secret keys is lost, the owner can still access the money if she has access to a second copy.
More sophisticated techniques can be used to achieve greater degrees of security. 
For example, the secret key can be divided in three parts in such a way that two out of the three parts are needed to access the money.
Secrets can also be kept in the brain, with no physical proof of the ownership of the money.

As cryptocurrencies are based on free libre software, anyone can create a new currency.
If a currency is used by a single individual it is not very useful.
The potential of networked cryptocurrencies materializes when they are used by communities.
The communities can create or agree in a currency for their internal exchanges.
As a result, the currency is impregnated by the values and ethical standards of the community.
In a sense, the value of cryptocurrencies is backed only by the community of users.

The people can choose which currency to use and which community to join.
In this choice, there is an implicit commitment to support certain values and principles.
By joining a particular community and currency, the user strengthens that particular community by offering productive force and services.
The currency is a commons resource used by the whole community.

The use of networked criptocurrencies evidences that flows of money are nothing more than flows of information.
Money transactions are simply communications among the involved parties.
The freedom to use different currencies is directly related to the freedom to communicate, which is popularly referred to as free speech.
The flows of money within a community are communications flows that strengthen the relation between the members of the community and, therefore, the community as a whole.

The value of the currency increases as the number of users and transactions increases.
Therefore, the members of a community have an incentive to devote some currency resources to attract new members to the community.
It is in their self-interest to make donations that increase the value of the currency.
In this regards, cryptocurrencies are a transition step towards a gift economy.

The gift (or reputation) economy is the kind of economy we find on the Internet, in which the participants offer software, articles, music, pictures and all kinds of digital goods for free.
This can be achieved using permissive licenses such as Creative Commons.
In the information age money is just information and wealth is measured in knowledge instead of material goods.
The progressive evolution towards a reputation economy is a natural step.

\section{Community-Centric Innovation}
\label{sec:other}

Throughout human history, organizational models have changed, transformed and evolved.
As new technologies develop, old models are progressively replaced by others that are more efficient in the context of the new technologies.
The invention of information processing and transmission tools multiplied the potential of community-centric models.
Initiatives embracing the community-centric model have an advantage compared to those attaching to old hierarchical approaches.

It is natural that we find salient examples of community-centric success in the field of information and communication technologies.
The Internet itself is probably the best example of a community-centric endeavour.
There is a large number of people fascinated by and contributing to the Internet.
The decision-making forums are open and anyone is free to contribute expertise.

Another example of the successful application of the community-centric model is the free libre open source movement.
Thousands of programmers affiliate to and contribute to this movement working as a self-managed community.
The software produced by this community often outperforms what have been developed using hierarchical models.
By making it possible for anyone to contribute, FLOSS projects have an advantage with respect to those that choose more restrictive models.

The Wikipedia is an Internet-based project that has also adopted the community-centric model.
The users are not mere consumers of information. 
On the contrary, the community of users is the one playing a central role by writing, revising and correcting the encyclopedical articles.
The community-centric model has allowed Wikipedia to grow in terms of number of articles, languages and quality.
In certain aspects, the community-centric encyclopedia has surpassed encyclopedias that were built using more traditional models.

There are also examples of success of community-centric models in hardware initiatives.
Both Arduino and BeagleBoard are open hardware projects that are aligned with community-centric innovation.
They both share all the information regarding the hardware with the community of users that, in turn, use this information to contribute to the project.

In the case of open hardware, the fact that all schematics and manufacturing details are publicly available on the Internet makes it very economical for a manufacturer to produce the goods.
The manufacturer can avoid all research and development costs and focus only in the fabrication.
As a consequence, open hardware devices can be found at very low prices in the market.
Lower prices translate to a greater community that can contribute to further improve the original product.

Decentralization, user empowerment and a focus on community-centric models seem to be a common denominator in the most exciting innovations in the area of information and communication technologies.

\section{Projects about Community Networks and Cryptocurrencies}
\label{sec:projects}

Community-centric innovation such as Community Networks and Networked Cryptocurrencies have attracted the attention of the research community.
In this section we detail three European projects that are related to community-centric innovation.

\subsection{Commons for Europe}
Commons for Europe is a project focused on stimulating the production of commons resources in the areas of applications and networks.
The branch of Commons for Europe related to application is called Code for Europe and invites cities to incorporate programmers to their teams.
These programmers bring community and collaborative dynamics to the host institution and write software that later can be shared with other cities.

The other branch of the Commons for Europe project is called Bottom up Broadband.
This branch is focused on understanding and documenting the working principles of community networks.

\subsection{Confine}
The Confine project \cite{braem2013crc} emphasizes the links between research and community networks.
The goal is to make it easier for researchers to run experiments in community networks.
It provides the tools for automatize and monitor the experiments, as well as the hardware distributed across the community network.

Community networks have grown to a size far beyond what can be reproduced in a lab.
Therefore, the most effective way to learn and understand this phenomena is by carrying out research directly on the community network.
It is a good opportunity for researchers to be in close contact with these networks that are leading the innovation in some areas of networking.
At the same time, the community network benefits from the expertise of the research community and from the documentation and dissemination efforts associated to the research process.

\subsection{D-Cent}
The area of cryptocurrency is covered by the D-Cent project.
This project is actually more ambitious and also considers decentralized social networks and tools for a more direct engagement in social decision making and deliberation.

The scope of this project also includes community-centric innovation, privacy and security, open knowledge, open data and open standards.

\section{Conclusions}
\label{sec:conclusion}

In some aspects of information and communication technologies, communities are  leading the path of innovation.
Community-centric innovation is characterized by peer-to-peer relations and openness.
Communities have also demonstrated their ability to take advantage of data networks and network effects.
In the case of community networks, a reinforcement loop is established in which a community deploys a network that, in its turn, strengthens the community.
Community networks have been with us for over a decade and have remained faithful to the original spirit of the Internet in terms of decentralization.
These networks are planned, financed, deployed and maintained by the same people that uses them.
They adopt self-management dynamics and can benefit from the combined knowledge of all the members of the community.

Despite being more time-consuming and less reliable for the user, they are still a valid alternative for some people.
The possibility of controlling the network and being participants instead of consumers are some of the reasons for joining a community network.
Another reason is that the idea of freedom is embedded in community networks.
Community networks can resist authoritarian censorship or Internet blackouts ordered by governments.
They are also extremely useful in case of a natural disaster when centralized infrastructures may collapse.

Networked cryptocurrencies are another example of community-centric innovation.
These currencies are developed by the communities as a result of open debate and discussion.
There are many similitudes with community networks.
The users of networked cryptocurrencies can contribute one node to the network to help in maintaining security.
The members of the community can also participate in many other ways offering products and services and using the cryptocurrency in their transactions.Networked cryptocurrencies

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

%\section{Conclusion}
%The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

% use section* for acknowledgement
%\section*{Acknowledgment}

%The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
%\ifCLASSOPTIONcaptionsoff
%  \newpage
%\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,my_bib}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Michael Shell}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{John Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%\begin{IEEEbiographynophoto}{Jane Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}